The resurgence of the linear optics \sihui{quantum} interferometer --- recent advances \& applications

### 2018-02-07 paper.tex
Touch up
Grammar check, style check, added figure for quantum scissors, bib
check.

Linear optics play a prominent role in quantum information processing \cite{bib:LovettKok}. Photons make fantastic `flying' qubits, and are readily used for quantum communication \cite{bib:Northup14} and quantum key distribution \cite{bib:Lo14}. In 2001, Knill, Laflamme and Milburn (KLM) showed that efficient quantum computing is possible using only linear optical components, that is single photons, beamsplitters, phase shifters and photon counting  \sihui{for spatially encoded qubits \cite{bib:KLM01}. At about the same time, Koashi, Yamamoto, and Imoto essentially came to the same result for polarization-encoded qubits \cite{Koashi01}.} Furthermore, an optical implementation of quantum protocols can potentially piggyback on existing communication infrastructure. Combined with the relative ease of transporting quantum states of light, as compared to say trapped ions or superconducting qubits, linear optics forms a powerful platform for quantum information processing.

@@

In the last decade, single photons have been produced mainly through \sihui{nonlinear optical processes know as spontaneous parametric down-conversion (SPDC) and spontaneous four-wave mixing (SFWM) \cite{bib:Eisaman11,bib:Xiong16,bib:Solntsev17}.} \sihui{Both SPDC and SFWM produce} single photons in correlated pairs that has been a staple source of entanglement. \sihui{However, SPDC uses up one pump photons per pair, while SFWM requires two.} Through developments in production, the number of high quality indistinguishable photons that can be simultaneously produced have increased steadily over the years. Single-photon pairs, once a novelty, are now run-of-the-mill business for quantum optical labs, and three or four photons have become commonplace \cite{bib:Spagnolo12, bib:Tillmann13, bib:Spring13,bib:Broome13}. Experiments using up to ten photons have been demonstrated \cite{bib:WangChen16, bib:Chen17}. We have also seen the meteoric rise of quantum dots in the role for producing single photons \cite{bib:Ding16,bib:WangHe16, bib:213601}. They are a scalable source of single photons which are highly indistinguishable, and are produced on demand--a significant advantage over SPDC which is an inherently probabilistic process.

Another approach to linear optics is to pack more features known as 
degrees of freedom onto a single photon, and to increase the degree 
of control over them. It is possible to manipulate degrees of
freedom like polarization, time-of-arrival, and orbital angular momentum \cite{bib:Tillmann2015,bib:Bozinovic2013, 
bib:Nicolas2014, bib:Humphreys2013, bib:Donohue2013}. Since KLM, new proposals 
for linear optical quantum computations (LOQC) have surfaced 
\cite{bib:Nielsen04,bib:BrowneRudolph05,bib:Gimeno-Segovia15,bib:Pant17}. 
\sihui {Some of them support new functionalities like secure delegated quantum 
computations \cite{bib:Barz12,bib:Fisher14}}. Others do not promise 
universality in their processing, but still performs tasks such as sampling 
\cite{bib:Aaronson11}, \sihui{encrypted quantum walks \cite{bib:Rohde_qw12}}, 
and secure quantum computations \cite{bib:Tan16} that might be hard or 
impossible with a classical system.  Such methods open up new ways to engineer 
quantum states and quantum operations, and expand the toolbox of baseline 
resources we have for implementing quantum computations, which could prove 
crucial in the race to achieve quantum supremacy.

We begin in Section \ref{sec:math} with an overview of the mathematical treatment of linear optical transformation on single photons, and in Section \ref{sec:encoding} with a description of some commonly used qubit encodings. Any unitary transformation acting on the spatial label of single photons can be efficiently decomposed into a network of beamsplitters and phase-shifters. We discuss this, and an extension to handle additional internal states, in Section \ref{sec:decomp}. For applying any linear optical circuit, one needs an accurate description of its unitary representation. A full quantum tomography is costly and time-consuming. Some practical methods have been devised to do this reconstruction without resorting to tomography, and these are described in Section \ref{sec:reconstruction}. In Section \ref{sec:state}, we recap the experimental breakthroughs in state preparation of single photons, and some entangled quantum states. This is followed by advances in time-bin based architecture for linear optical networks, and photodectors in Sections \ref{sec:timebins} and \ref{sec:detector} respectively. Last, but not least, we review applications for linear optical interferometry in LOQC, {\sc bosonsampling}, \sihui{ quantum metrology, and quantum repeater networks}.

@@

A single photon in a quantum interferometer is described by its creation and annihilation operators, $\adagh_{j}$ and $\hat{a}_j$ respectively, where $j$ is the mode label of the interferometer. These operators satisfy the bosonic comutation relationship $[\ah_{j},\adagh_{k}]=\delta_{j,k}$. The action of a $2d$-port linear optical interferometer that has an equal number of input and output ports is expressed as an application of unitary operations on the creation operators,
\begin{align}
b_i^{\dag}=\sum_{j=1}^d U_{ij}a_j^\dag \ ,
\end{align}
where $a_j^\dag$ and $b_i^\dag$ are the creation operators of a single input and output photon in the $j$-th and $i$-th modes respectively, and $U\in SU(d)$. All such transformations can be expressed as sequences of beamsplitters and phase-shifters \cite{bib:Reck1994} (see Section \ref{sec:decomp}). By convention, the interferometer is assumed to act only on the spatial mode of the input state.

@@

Control of indistinguishability of these photons may enable future applications. Mathematical methods have been developed to deal with partial distinguishabilities among interfering photons, including those using group theory \cite{bib:Tan2013,bib:deGuise2014,bib:deGuise2015}, and quantum-to-classical transitions \cite{bib:Ra2013}. Recently, by controlling multiple degrees of freedom of single photons, genuine three-photon interference that has no independent entanglement between any two subpairs of photons was demonstrated  \cite{bib:Agne17,bib:Menssen17}.

@@

In dual-rail encoding, a single photon encodes a logical qubit as a superposition across two spatial modes,
\begin{align}
\ket\psi_L = \alpha\ket{1,0} + \beta\ket{0,1},
\end{align}
where $\ket{i,j}$ is a two-mode state with $i$ ($j$) photons in the first (second) spatial mode. Using this encoding, phase-shifters and beamsplitters between the two spatial modes implement arbitrary single-qubit operations. Converting between polarisation- and dual-rail-encoding is trivial using polarising beamsplitters, which separate horizontal and vertical components into distinct spatial modes, or vice-versa.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.4\columnwidth]{reck}
\caption{Efficient decomposition of arbitrary linear optics 
networks into a sequence of beamsplitters and phase-shifters. 
\sihui{This setup describes a generalized multi-port Mach-
Zehnder interferometer}} \label{fig:Reck}	
\end{figure}

@@

State tomography is achieved by measuring expectation values in a complete basis for the state Hilbert space. For one qubit, for example, a complete basis is the Pauli basis, which wields the completeness relation,}
\begin{align}
\hat\rho = \text{tr}({\hat\rho}) + \text{tr}(\hat{X}\hat{\rho})\hat{X} + \text{tr}({\hat{Y}\hat\rho})\hat{Y} + \text{tr}(\hat{Z}{\hat\rho})\hat{Z}.
\end{align}
\peter{These traces correspond exactly to the expectation values in the respective bases, which are easily determined using separable measurements. \sihui{Sometimes, it is possible to be able to do a tomography from an incomplete number of measurements, {\it i.e.} a number that is less than the number of expectation values, and it is possible using compressed sensing provided some sparsity conditions are met \cite{bib:GLFBE2010,bib:Titchener16,bib:Oren16}.} Process tomography employs state tomography as a primitive, applying an equivalent completeness relation, but across the space of quantum processes, which may be decomposed into a basis input to output states. For a single qubit process this may be represented as,}
\begin{align} \label{eq:process_matrix}
\mathcal{E}(\hat\rho) = \sum_{i,j=1}^4 \chi_{i,j} \hat{\sigma}_i\hat\rho\,\hat{\sigma}_j^\dag.
\end{align}
\peter{which generalises to an arbitrary number of qubits. Here $\chi$ is the process matrix in the Pauli basis.}
%\cite{bib:Mitchell03,bib:Obrien04,bib:Lobino08,bib:Saleh11}.

@@

The GHZ state is also known as the NOON state, because it is written as
\begin{align}
\ket{\psi^{\rm (NOON)}_n}=\frac{1}{\sqrt{2}}(\ket{n,0}+\ket{0,n})\ ,
\end{align}
in the second quantization representation. This notation is commonly used in quantum metrological applications.

@@

\sihui{
\subsection{Quantum state engineering based on the multi-port Mach-Zehnder interferometer}
Linear optical inteferometers have been used for quantum state engineering. One well-known example is that of optical state truncation, sometimes also referred to as quantum scissors, and it happens by way of a process known as projection synthesis [figure for this?].  Projection synthesis works by projecting an input state onto an output state by putting the former through a multiport interferometer, and then postselecting on a certain photon number output. Optical state truncation is a use of projection synthesis to use a continuous variable state represented in the Fock-state basis, $\ket{\phi}=\sum_{n=0}^\infty c_n\ket{n}$, to produce the normalized form of $\sum_{n=0}^d c_n\ket{n}$. The idea of projection synthesis was first conceived for the special case of $d=2$ \cite{bib:Pegg98, bib:Barnett99}, and later extended to $d=3$ \cite{bib:Koniorczyk2000}, and arbitrarily large $d$ \cite{bib:Miranowicz07, bib:Goyal13}. Notably, the scheme proposed in \cite{bib:Goyal13} uses multimode single photons and repeated uses of $d=2$ quantum scissors to extend the depth of truncation. 
}

@@

Using time-bin encoding of optical qubits, the obvious question is how to perform operations upon them. In \cite{bib:Motes14}, a dual-loop architecture was presented for implementing arbitrary passive linear-optics on photonic pulse-trains with time-bin separation $\tau$, shown in Fig.~\ref{fig:FL_arch}. The inner loop has lengtxh exactly $\tau$, while the outer one has length $>n\tau$ ($n$ is the 
number of optical modes). The architecture is controlled via three 
dynamically controlled beamsplitters. The first and last need only be on/off switches, whose sole purpose is to couple in the prepared pulse-train, 
keep it within the outer loop for the required duration, and then couple 
out of the outer loop, yielding the transformed pulse-train. The central 
beamsplitter must be able to implement arbitrary classically-controlled 
beamsplitter operations.

@@

The above description applies to passive linear optics, which is sufficient for protocols such as {\sc BosonSampling}, but insufficient for universal optical quantum computation, which requires the addition of ancillary states, and measurement with fast-feedforward. To address this, it was shown in \cite{Rohde15} that by dynamically preparing ancillary pulse-trains (from the already-existing source), classically controlled by time-resolved measurements at the output, and changing the switching sequence, we can effectively couple in and out arbitrary subsets of the optical modes, enabling partial measurements to be implemented.

@@

Broadly speaking, there are two main classes of photodetectors: non-photon-number-resolving (also referred to as bucket, or on/off) detectors, and photon-number-resolving (PNR) detectors. As the names suggest, non-PNR detectors only detect the presence or absence of photons in an incident beam, while PNR detectors are able to measure its number of photons. The outcomes of a PNR detector are described by the measurement projectors,
\begin{align}
\hat\Pi_n &= \ket{n}\bra{n}, \ n=0,1,\ldots \ ,
\end{align}
while those of the non-photon-number-resolving detectors are given by
\begin{align}
\hat\Pi_\text{off} = \hat\Pi_0, \ {\rm and} \ \hat\Pi_\text{on} = \hat{I} - \hat\Pi_\text{on}\ .
\end{align}
\sihui{Theoretical modeling of these photodetectors can be made to include realistic effects such as detector finite efficiency, finite-number resolution, dark counts, and crosstalk. These effects are typically described as projections onto mixed states \cite{bib:BARNETT1998}, and their outcomes are represented by positive-operator valued measures. Such modeling allows us to analyze these realistic effects on various quantum information protocols, see for {\it e.g.~}\cite{bib:Ozdemir2001,bib:Sperling2014,bib:Tan2016}.}

@@

PNR detectors are a wonderful addition to the \sihui{linear optical} toolbox. A notable example is the superconducting transition edge sensor (TES) which has an extremely high quantum efficiency. The principle it operates on is very similar to the SNSPD, but with a much higher sensitivity that allows it to detect the energy of single photons. \sihui{They have low dark count rates, however, they are also quite slow, have low maximum counting rates, and, like the SNSPD detectors, are costly to operate due to their very low operating temperatures.} The visible light photon counter (VLPC) is another kind of PNR detector. The way VLPC works is in principle very similar to the APD, but with an additional layer of silicon that is lightly doped with arsenic (As). Owing to the presence of the As, a single photon absorption event always create an electrical signal that is always of the same magnitude. Thus, the output electrical signal is just proportional to the number of detected photons. \sihui{The VLPC can achieve a high quantum efficiency, in fact almost as high as that of the TES, and has a distinct advantage over the TES of not requiring superconducting temperatures. However, they do suffer from high dark count rate, and low speeds.}
Last, PNR detectors can be created by connecting many non-PNR detectors in parallel to one another, and summing the signals at the output. For instance, like in the theoretical proposal of \cite{bib:Sperling12}. Such an approach has been successfully demonstrated with SNSPD \cite{bib:Divochiy08}, and APDs \cite{bib:Kalashnikov12,bib:Chrapkiewicz14,bib:Heilmanndc15}. \sihui{Some characteristic experimental values for the detector efficiency, timing jitter, dark count rates, and maximum photon number of different photodetectors are given in Table \ref{table:PNR}.}

\begin{table}[t]\label{table:PNR}
\centering
\begin{threeparttable}
\caption{Comparison of performance between some photodetectors. In part adapted from \cite{bib:Eisaman11}.}\label{table:PNR}
\begin{tabular}{| >{\centering\arraybackslash}m{2cm} | >{\centering\arraybackslash}m{3cm} | >{\centering\arraybackslash}m{2cm} |>{\centering\arraybackslash}m{2cm} |>{\centering\arraybackslash}m{2cm} |} 
 \hline 
\vspace{0.8mm} Detector type & detector efficiency (\%) & timing jitter ($ns$) & dark counts ($1/s$) & maximum photon number \\ \hhline{|=|=|=|=|=|}
&&&& \\
APD \cite{bib:Webster2012} & 72@560$nm$ & 0.09 & 18 & 1\tnote{a} \\
&&&& \\
SNSPD \cite{bib:Takesue07} & 0.7@1550$nm$ & 0.06 & 10 & 1\tnote{a} \\
&&&&\\
  VLPC \cite{bib:Waks03}   & 85@543$nm$  & 2 & 20000 & 5 \\
  & & & & \\
  TES \cite{bib:Lita08} & 95@1556 & 100 & - & 7\\
 TES \cite{bib:MillerNam03} & 20@1310$nm$  & 100 & $<10^{-3}$ & 15 \\
  & 20@1550$nm$ & & & \\
  TES \cite{bib:Gerrits12} & 89@1550$nm$ & - & - & 1000 \\
  &&&& \\
 Parallel SNSPD \cite{bib:Divochiy08} & 2@1300 & 0.05 & 0.15 & 4 \\
 &&&&\\
 Series SNSPD \cite{bib:Mattioli16} & 0.5@1310 & 0.116 & - & 24\\
&&&& \\
 \hline
\end{tabular}
\begin{tablenotes}
\item[a] Has no photon-number-resolving capabilities.
\end{tablenotes}
\end{threeparttable}
\end{table}
%\footnote{Has no photon-number-resolving capabilities.}
\section{Optical switching}

\peter{
Some of the protocols thus far require switching between different optical modes. \sihui{One way to achieve this experimentally for time-bin qubits is} via a Mach-Zehnder interferometer with a dynamic, classically-controlled phase-shift in one path \cite{bib:Takesue14}. The variable phase may be implemented in a number of ways using various types of phase modulators, such as electro-optic or acousto-optic modulators. With this construction setting the phase to $\phi=0$ performs the identity operation, whereas for $\phi=\pi$ it performs a swap gate. This is shown in Fig. \ref{figMZ}. \sihui{Ref.~\cite{bib:Takesue14} reports a switching time of 100 picoseconds for this type of switch.}
}

@@

 \sihui{It is also possible to perform switching of photons by first converting temporal and spatial labels of the photon to a polarization one, and then perform a fast polarization switch before converting it back \cite{bib:HAK11,bib:Humphreys13}. Such switches are able to achieve a polarization switch within a 10 picoseconds time window. Other ways rely on conditional scattering of signal photons on ultrafast timescales, usually involving coupling between field modes and strongly correlated systems \cite{bib:Volz2012, bib:Bose2012, bib:Dietrich16}. The typical speed for this type of switching is of order of picoseconds \cite{bib:Dietrich16}.
 }

\sihui{This concept for switching is widely used in current labs, however it is very slow using present-day technology. Although switching on its own can be completed in a timespan of the order of picoseconds, the process is slowed by the time needed for its control electronic. This is especially crucial in switching for the purposes of fast-feedforward that is required, for example, for universal optical quantum computing. In this application, we would require total switching times, including time for control electronics to respond, on the order of nanoseconds, {\it i.e.~}faster than the time to the destination of the light it needs to switch, and this is still not possible. Presently this remains a daunting challenge, requiring much further research, with quantum memories being a candidate for loosening the switching time limitations.
}

@@

\section{Applications for linear optics interferometry}

@@

Linear optical interferometry has become a leading contender for the implementation of quantum information processing protocols. The preparation, manipulation and measurement of photonic quantum information has become mainstream and widely employed, with impressive experimental accuracies.

>>>>

Linear optics play a prominent role in quantum information processing \cite{bib:LovettKok}. Photons make fantastic `flying' qubits, and are readily used for quantum communication \cite{bib:Northup14} and quantum key distribution \cite{bib:Lo14}. In 2001, Knill, Laflamme and Milburn (KLM) showed that efficient quantum computing is possible using only linear optical components, that is single photons, beamsplitters, phase shifters and photon counting  \sihui{for spatially encoded qubits \cite{bib:KLM01}. At about the same time, Koashi, Yamamoto, and Imoto (KYI) essentially came to the same result for polarization-encoded qubits \cite{Koashi01}.} Furthermore, an optical implementation of quantum protocols can potentially piggyback on existing communication infrastructure. Combined with the relative ease of transporting quantum states of light, as compared to say trapped ions or superconducting qubits, linear optics forms a powerful platform for quantum information processing.

@@

In the last decade, single photons have been produced mainly through \sihui{nonlinear optical processes know as spontaneous parametric down-conversion (SPDC) and spontaneous four-wave mixing (SFWM) \cite{bib:Eisaman11,bib:Xiong16,bib:Solntsev17}.} \sihui{Both SPDC and SFWM produce} single photons in correlated pairs that has been a staple source of entanglement. \sihui{However, SPDC uses up one pump photons per pair, while SFWM requires two.} Through developments in production, the number of high quality indistinguishable photons that can be simultaneously produced have increased steadily over the years. Single-photon pairs, once a novelty, are now run-of-the-mill business for quantum optical \sihui{laboratories}, and three or four photons have become commonplace \cite{bib:Spagnolo12, bib:Tillmann13, bib:Spring13,bib:Broome13, bib:Crespi13}. Experiments using up to ten photons have been demonstrated \cite{bib:WangChen16, bib:Chen17}. We have also seen the meteoric rise of quantum dots in the role for producing single photons \cite{bib:Ding16,bib:WangHe16, bib:213601}. They are a scalable source of single photons which are highly indistinguishable, and are produced on demand--a significant advantage over \sihui{nonlinear optical processes which are inherently probabilistic.}

Another approach to linear optics is to pack more features known as 
degrees of freedom onto a single photon, and to increase the degree 
of control over them. It is possible to manipulate degrees of
freedom like polarization, time-of-arrival, and orbital angular momentum \cite{bib:Tillmann2015,bib:Bozinovic2013, 
bib:Nicolas2014, bib:Humphreys2013, bib:Donohue2013}. Since KLM \sihui{and KYI}, new proposals 
for linear optical quantum computations (LOQC) have surfaced 
\cite{bib:Nielsen04,bib:BrowneRudolph05,bib:Gimeno-Segovia15,bib:Pant17}. 
\sihui {Some of them support new functionalities like secure delegated quantum 
computations \cite{bib:Barz12,bib:Fisher14, bib:Tan16}}. Others do not promise universality in their processing, but still performs tasks such as sampling \cite{bib:Aaronson11}, and \sihui{encrypted quantum walks \cite{bib:Rohde_qw12}} that might be hard or 
impossible with a classical system.  Such methods open up new ways to engineer 
quantum states and quantum operations, and expand the toolbox of baseline 
resources we have for implementing quantum computations, which could prove 
crucial in the race to achieve quantum supremacy.

We begin in Section \ref{sec:math} with an overview of the mathematical treatment of linear optical transformation on single photons, and in Section \ref{sec:encoding} with a description of some commonly used qubit encodings. Any unitary transformation acting on the spatial label of single photons can be efficiently decomposed into a network of beamsplitters and phase-shifters. We discuss this, and an extension to handle additional internal states, in Section \ref{sec:decomp}. For applying any linear optical circuit, one needs an accurate description of its unitary representation. A full quantum tomography is costly and time-consuming. Some practical methods have been devised to do this reconstruction without resorting to tomography, and these are described in Section \ref{sec:reconstruction} \sihui{while quantum tomography is briefly discussed in Section \ref{secQPT}}. In Section \ref{sec:state}, we recap the experimental breakthroughs in state preparation of single photons, and some entangled quantum states. This is followed by advances in time-bin based architecture for linear optical networks, \sihui{photodectors, and optical switches in Sections \ref{sec:timebins}, \ref{sec:detector}, and \ref{sec:switch} respectively}. Last, but not least, we review applications for linear optical interferometry in \sihui{Section \ref{sec:applications}}.

@@

A single photon in a quantum interferometer is described by its creation and annihilation operators, $\adagh_{j}$ and $\hat{a}_j$ respectively, where $j$ is the mode label of the interferometer. These operators satisfy the bosonic comutation relationship $[\ah_{j},\adagh_{k}]=\delta_{j,k}$. The action of a $2d$-port linear optical interferometer that has an equal number of input and output ports is expressed as an application of unitary operations on the creation operators,
\begin{align}
b_i^{\dag}=\sum_{j=1}^d U_{ij}a_j^\dag \ ,
\end{align}
where $a_j^\dag$ and $b_i^\dag$ are the creation operators of a single input and output photon in the $j$-th and $i$-th modes respectively, and $U\in {\rm SU}(d)$. All such transformations can be expressed as sequences of beamsplitters and phase-shifters \cite{bib:Reck1994} (see Section \ref{sec:decomp}). By convention, the interferometer is assumed to act only on the spatial mode of the input state.

@@

Control of indistinguishability of these photons may enable future applications. Mathematical methods have been developed to deal with partial distinguishabilities among interfering photons, including those using group theory \cite{bib:Tan2013,bib:deGuise2014,bib:Shchesnovich15,bib:deGuise2015}, and quantum-to-classical transitions \cite{bib:Ra2013}. Recently, by controlling multiple degrees of freedom of single photons, genuine three-photon interference that has no independent entanglement between any two subpairs of photons was demonstrated  \cite{bib:Agne17,bib:Menssen17}.

@@

In dual-rail encoding, a single photon encodes a logical qubit as a superposition across two spatial modes,
\begin{align}\label{eq:dualrail}
\ket\psi_L = \alpha\ket{1,0} + \beta\ket{0,1},
\end{align}
where $\ket{i,j}$ is a two-mode state with $i$ ($j$) photons in the first (second) spatial mode. Using this encoding, phase-shifters and beamsplitters between the two spatial modes implement arbitrary single-qubit operations. Converting between polarisation- and dual-rail-encoding is trivial using polarising beamsplitters, which separate horizontal and vertical components into distinct spatial modes, or vice-versa.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.55\columnwidth]{reck}
\caption{Efficient decomposition of arbitrary linear optics 
networks into a sequence of beamsplitters and phase-shifters. 
\sihui{This setup describes a generalized $n\times n$ multi-port Mach-Zehnder interferometer. Adapted from \cite{bib:Reck1994}.}} \label{fig:Reck}	
\end{figure}

@@

State tomography is achieved by measuring expectation values in a complete basis for the state Hilbert space. For one qubit, for example, a complete basis is the Pauli basis, which wields the completeness relation,}
\begin{align}
\hat\rho = \text{tr}({\hat\rho}) + \text{tr}(\hat{X}\hat{\rho})\hat{X} + \text{tr}({\hat{Y}\hat\rho})\hat{Y} + \text{tr}(\hat{Z}{\hat\rho})\hat{Z}.
\end{align}
\peter{These traces correspond exactly to the expectation values in the respective bases, which are easily determined using separable measurements. \sihui{Sometimes, it is possible to be able to do a tomography from an incomplete number of measurements, {\it i.e.} a number that is less than the number of expectation values, and it is possible using compressed sensing provided some sparsity conditions are met \cite{bib:GLFBE2010,bib:Titchener16,bib:Oren16}. In cases where one only wishes to know if a given state is close to a presumed state, one can perform a certification test instead of state tomography \cite{bib:Aolita15}.}

Process tomography employs state tomography as a primitive, applying an equivalent completeness relation, but across the space of quantum processes, which may be decomposed into a basis input to output states. For a single qubit process this may be represented as,}
\begin{align} \label{eq:process_matrix}
\mathcal{E}(\hat\rho) = \sum_{i,j=1}^4 \chi_{i,j} \hat{\sigma}_i\hat\rho\,\hat{\sigma}_j^\dag.
\end{align}
\peter{which generalises to an arbitrary number of qubits. Here $\chi$ is the process matrix in the Pauli basis.}
%\cite{bib:Mitchell03,bib:Obrien04,bib:Lobino08,bib:Saleh11}.

@@

The GHZ state is also known as the NOON state, because it is written as
\begin{align}
\ket{\psi^{\rm (NOON)}_n}=\frac{1}{\sqrt{2}}(\ket{n,0}+\ket{0,n})\ ,
\end{align}
in the second quantization representation, \sihui{and the basis is the same as that for eq.~(\ref{eq:dualrail})}. This notation is commonly used in quantum metrological applications.

@@

\sihui{
\subsection{Quantum state engineering based on the multi-port Mach-Zehnder interferometer}
Linear optical inteferometers have been used for quantum state engineering. One well-known example is that of optical state truncation, sometimes also referred to as quantum scissors, and it happens by way of a process known as projection synthesis.  Projection synthesis works by projecting an input state onto an output state by putting the former through a multiport interferometer, and then postselecting on a certain photon number output. Optical state truncation is a use of projection synthesis to use a continuous variable state represented in the Fock-state basis, $\ket{\phi}=\sum_{n=0}^\infty c_n\ket{n}$, to produce the normalized form of $\sum_{n=0}^d c_n\ket{n}$. An example of the interferometer used for optical state truncation for the case $d=1$ is shown in Fig.~\ref{fig:qscissors}. The idea of projection synthesis was first conceived for the special case of $d=2$ \cite{bib:Pegg98, bib:Barnett99}, and later extended to $d=3$ \cite{bib:Koniorczyk2000}, and arbitrarily large $d$ \cite{bib:Miranowicz07, bib:Goyal13}. Notably, the scheme proposed in \cite{bib:Goyal13} uses multimode single photons and repeated uses of $d=2$ quantum scissors to extend the depth of truncation. 
}

\begin{figure}[!htb]
\centering
\includegraphics[width=0.3\columnwidth]{qscissors}
\caption{Figure showing optical state truncation for $d=1$. The input state is $\ket{\phi}$, and when the photodetectors $D_1$ and $D_2$ register one and zero photons respectively, the output state is $\ket{\phi}_{\rm out}=c_0\ket{0}+c_1\ket{1}$ with an appropriate normalization factor. Adapted from \cite{bib:Pegg98}.} \label{fig:qscissors}	
\end{figure}

@@

\sihui{When} using time-bin encoding of optical qubits, the obvious question is\sihui{: how do we perform single-qubit and two-qubit} operations upon them. In \cite{bib:Motes14}, a dual-loop architecture was presented for implementing arbitrary passive linear-optics on photonic pulse-trains with time-bin separation $\tau$, shown in Fig.~\ref{fig:FL_arch}. The inner loop has length exactly $\tau$, while the outer one has length $>n\tau$ ($n$ is the 
number of optical modes). The architecture is controlled via three 
dynamically controlled beamsplitters. The first and last need only be on/off switches, whose sole purpose is to couple in the prepared pulse-train, 
keep it within the outer loop for the required duration, and then couple 
out of the outer loop, yielding the transformed pulse-train. The central 
beamsplitter must be able to implement arbitrary classically-controlled 
beamsplitter operations.

@@

The above description applies to passive linear optics, which is sufficient for protocols such as {\sc BosonSampling}, but insufficient for universal optical quantum computation, which requires the addition of ancillary states, and measurement with fast-feedforward. To address this, it was shown in \cite{Rohde15} that \sihui{partial measurements can be implemented by coupling in and out arbitrary subsets of the optical modes. The coupling is in turn achieved via dynamically preparing ancillary pulse-trains (using the already existing source) that are classically controlled by time-resolved measurements at the output, and changing the switching sequence conditional on these measurement outcomes.}

@@

\begin{table}[t]\label{table:PNR}
\centering
\begin{threeparttable}
\caption{Comparison of performance between some photodetectors. In part adapted from \cite{bib:Eisaman11}.}\label{table:PNR}
\begin{tabular}{| >{\centering\arraybackslash}m{2cm} | >{\centering\arraybackslash}m{3cm} | >{\centering\arraybackslash}m{2cm} |>{\centering\arraybackslash}m{2cm} |>{\centering\arraybackslash}m{2cm} |} 
 \hline 
\vspace{0.8mm} Detector type & detector efficiency (\%) & timing jitter ($ns$) & dark counts ($1/s$) & maximum photon number \\ \hhline{|=|=|=|=|=|}
&&&& \\
APD \cite{bib:Webster2012} & 72@560$nm$ & 0.09 & 18 & 1\tnote{a} \\
&&&& \\
SNSPD \cite{bib:Takesue07} & 0.7@1550$nm$ & 0.06 & 10 & 1\tnote{a} \\
&&&&\\
  VLPC \cite{bib:Waks03}   & 85@543$nm$  & 2 & 20 000 & 5 \\
  & & & & \\
  TES \cite{bib:Lita08} & 95@1556$nm$ & 100 & - & 7\\
 TES \cite{bib:MillerNam03} & 20@1310$nm$  & 100 & $<10^{-3}$ & 15 \\
  & 20@1550$nm$ & & & \\
  TES \cite{bib:Gerrits12} & 89@1550$nm$ & - & - & 1000 \\
  &&&& \\
 Parallel SNSPD \cite{bib:Divochiy08} & 2@1300$nm$ & 0.05 & 0.15 & 4 \\
 &&&&\\
 Series SNSPD \cite{bib:Mattioli16} & 0.5@1310$nm$ & 0.116 & - & 24\\
 &&&&\\
 MPPC \cite{bib:Afek09} & 50@400$nm$ & 20 & 115 000 & 10 \\
 & 8@800$nm$&&& \\
&&&& \\
 \hline
\end{tabular}
\begin{tablenotes}
\item[a] Has no photon-number-resolving capabilities.
\end{tablenotes}
\end{threeparttable}
\end{table}

Broadly speaking, there are two main classes of photodetectors: non-photon-number-resolving (also referred to as bucket, or on/off) detectors, and photon-number-resolving (PNR) detectors. As the names suggest, non-PNR detectors only detect the presence or absence of photons in an incident beam, while PNR detectors are able to measure its number of photons. The outcomes of a PNR detector are described by the measurement projectors,
\begin{align}
\hat\Pi_n &= \ket{n}\bra{n}, \ n=0,1,\ldots \ ,
\end{align}
while those of the non-photon-number-resolving detectors are given by
\begin{align}
\hat\Pi_\text{off} = \hat\Pi_0, \ {\rm and} \ \hat\Pi_\text{on} = \hat{I} - \hat\Pi_\text{on}\ .
\end{align}
\sihui{Theoretical modeling of these photodetectors can be made to include realistic effects such as detector finite efficiency, finite-number resolution, dark counts, and crosstalk \cite{bib:LeeYurtsever04, bib:Ramilli10,bib:Kalashnikov12}. These effects are typically described as projections onto mixed states \cite{bib:BARNETT1998}, and their outcomes are represented by positive-operator valued measures. Such modeling allows us to analyze these realistic effects on various quantum information protocols, see for {\it e.g.~}\cite{bib:Ozdemir2001,bib:Sperling2014, bib:Matthews16}, and update our characterization of non-classicality \cite{bib:Sperlingbinomial12,bib:Kalashnikov12b, bib:Sperling15,bib:Heilmanndc15, bib:Tan2016}.}

@@

PNR detectors are a wonderful addition to the \sihui{linear optical} toolbox. A notable example is the superconducting transition edge sensor (TES) which has an extremely high quantum efficiency. The principle it operates on is very similar to the SNSPD, but with a much higher sensitivity that allows it to detect the energy of single photons. \sihui{They have low dark count rates, however, they are also quite slow, have low maximum counting rates, and, like the SNSPD detectors, are costly to operate due to their very low operating temperatures.} The visible light photon counter (VLPC) is another kind of PNR detector. The way VLPC works is in principle very similar to the APD, but with an additional layer of silicon that is lightly doped with arsenic (As). Owing to the presence of the As, a single photon absorption event always create an electrical signal that is always of the same magnitude. Thus, the output electrical signal is just proportional to the number of detected photons. \sihui{The VLPC can achieve a high quantum efficiency, in fact almost as high as that of the TES, and has a distinct advantage over the TES of not requiring superconducting temperatures. However, they do suffer from high dark count rate, and low speeds.}
Last, PNR detectors can be created by connecting many non-PNR detectors in parallel to one another, and summing the signals at the output. For instance, like in the theoretical proposal of \cite{bib:Sperling12}. Such an approach has been successfully demonstrated with SNSPD \cite{bib:Divochiy08}, and spatially multiplexed APDs known as multi-pixel photon counters (MPPCs) \cite{bib:Afek09, bib:Chrapkiewicz14}. \sihui{Characteristic experimental values for the detector efficiency, timing jitter, dark count rates, and maximum photon number of some types of photodetectors are given in Table \ref{table:PNR}.}

\section{Optical switching}\label{sec:switch}

\peter{
Some of the protocols thus far require switching between different optical modes. \sihui{One way to achieve this experimentally for time-bin qubits uses} a Mach-Zehnder interferometer with a dynamic, classically-controlled phase-shift in one path \cite{bib:Takesue14} (see Fig.~\ref{figMZ}). The variable phase may be implemented in a number of ways using various types of phase modulators, such as electro-optic or acousto-optic modulators. With this construction setting the phase to $\phi=0$ performs the identity operation, whereas for $\phi=\pi$ it performs a swap gate.\sihui{Ref.~\cite{bib:Takesue14} reports a switching time of 100 picoseconds for this type of switch.}
}

@@

 \sihui{Another way to perform switching is to first convert temporal and spatial labels of the photon to a polarization one, and then perform a fast polarization switch before converting it back \cite{bib:HAK11,bib:Humphreys13}. Such switches are able to achieve a polarization switch within a 10 picoseconds time window. Other ways rely on conditional scattering of signal photons on ultrafast timescales, usually involving coupling between field modes and strongly correlated systems \cite{bib:Volz2012, bib:Bose2012, bib:Dietrich16}. The typical speed for this type of switching is of order of picoseconds \cite{bib:Dietrich16}.
 }

\sihui{This concept for switching is widely used in current laboratories, however it is very slow using present-day technology. Although switching on its own can be completed in a timespan of the order of picoseconds, the process is slowed by the response time needed for its electronic control. Overcoming this hurdle is especially crucial in switching for fast-feedforward that is required, for example, in universal optical quantum computing. In this application, we would require total switching times, including time for control electronics to respond, on the order of nanoseconds, {\it i.e.~}faster than the time to the destination of the light it needs to switch, and this is still not possible. Presently this remains a daunting challenge, requiring much further research, with quantum memories being a candidate for loosening the switching time limitations.
}

@@

\section{Applications for linear optics interferometry}\label{sec:applications}

@@

Linear optical interferometry has become a leading contender for the implementation of quantum information processing protocols. The preparation, manipulation and measurement of photonic quantum information \sihui{is already} mainstream and widely employed, with impressive experimental accuracies.
